---
title: "inspectDF: R package for getting inspected rules as data.frame"
author: Satoshi Kato (@katokohaku)
output: 
  md_document:
    variant: markdown_github
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE
)
```

# Overview

Getting inspected rules as data.frame.

## example

Just use `inspectDF()` instead of `arules::inspect()` after doing `arules::apriori()`.

```{r example.inspectDF}
require(dplyr)
require(arules)
require(inspectDF)

data(Groceries)
params <- list(confidence=0.001, support=0.001, maxlen=7, minlen=2)
glo.apriori <- apriori(Groceries, parameter = params, control = list(verbose=FALSE))
print(glo.apriori)

glo.inspectDF  <- inspectDF(glo.apriori)
```

**InspectDF** also provides a plot.igraph wrapper utility.

```{r example.plot}
set.seed(0)
glo.inspectDF %>% 
  arrange(support, confidence) %>%
  head(60) %>% 
  plotRuleGraph()

```

After `arules::apriori()`, just use `inspectDF()` instead of `arules::inspect()`. InspectDF has a good affinity with tidy schemes, such as `dplyr::arrange()` or `dplyr::filter()`.

```{r example.tidy}
set.seed(0)
glo.inspectDF  <- inspectDF(glo.apriori)
glo.inspectDF %>% 
  filter(n == 3) %>% 
  arrange(support, confidence) %>%
  head() %>% 
  knitr::kable()

```

## Installation

You can install the **inspectDF** package from [GitHub](https://github.com/katokohaku/inspectDF).

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("katokohaku/inspectDF")
```

The source code for **inspectDF** package is available on GitHub at

- https://github.com/katokohaku/inspectDF.

## Motivation

Usually, we do `inspect()` to enumrate rules after `arules::apriori()`.

Of cource, we could get data.frame object as side effect of `cat()` in `inspect()`. However, it can't be done quietly. It is noisy when using in function ow with a lot of rules.

In addition, arules package privides several utilities such as sort(), subset() and etc. But if rules were provided as data.frame, we can explore them as tidy data.

### Use case

InspectDF has a good affinity with tidy schemes, such as `dplyr::arrange()` or `dplyr::filter()`. E.g. rules with specific item(s) can be extracted using 

```{r usecase.filter}
require(stringr)
rules.lhs  <- glo.inspectDF %>% 
  filter(str_detect(LHS, pattern = "yogurt|sausage")) %>%
  arrange(confidence, lift) %>%
  filter(n > 1) %>% 
  head()
rules.lhs %>% knitr::kable()
```
 
By default, rule strings are split by separater `","`. But, items sometimes contain separater characters ***e.g. [IUPAC of DHA](https://pubchem.ncbi.nlm.nih.gov/compound/Docosahexaenoic_acid#section=IUPAC-Name)***. In such case, user can change rule-separater freely `sep = string`.
 
```{r usecase.sep}
glo.apriori %>% 
  inspectDF(sep = "###") %>% 
  filter(n >3) %>% 
  select(1:3) %>% 
  head() %>% 
  knitr::kable()
```

Similar to original plot.rules with igraph in arules package, each rule size represents **support value**. This size can be adjusted by `adujust.support.size` in plot functions. 

```{r usecase.adjust.rule}
require(stringr)
rules.lhs  <- glo.inspectDF %>% 
  filter(str_detect(RHS, pattern = "yogurt")) %>%
  arrange(confidence, lift) %>%
  filter(n > 1) %>% 
  head(15)

par(mfrow = c(1,2))
set.seed(0)
rules.lhs %>% plotRuleGraph(label = "default size")
set.seed(0)
rules.lhs %>% plotRuleGraph(label = "adjusted rule size", 
                            adujust.support.size = 4000)
par(mfrow = c(1,1))

```

